Improving Adaptive Seamless Designs through Bayesian optimization

### 2021-04-12 response_to_the_reviewers.md

### General

1) Power is used as a key performance metric, yet it lacks an explicit definition. The problem, of course, is that power is slightly more complicated in a multi-stage design with multiple arms (and therefore multiple hypotheses to be tested) than it is in a simple comparison of two groups. Given the scenario described in the motivating example, I would have assumed that power was defined as the probability of at least one (any) treatment group being demonstrated to be ‘better’ than control i.e. a ‘disjunctive’ or ‘minimal’ power definition (cf. Senn & Bretz 2007). However, on p. 10 it is stated that “the vector of treatment numbers for determining power counts rejections of 3 or 4 hypotheses”, implying a power definition more similar to ‘conjunctive’ or ‘maximal’ power.

Reference: Senn S, Bretz F (2007) Power and sample size when multiple endpoints are considered. Pharmaceutical Statistics, 6(3), 161-170.

> As in Friede et. al. (2020), where our motivating example was introduced, we count the rejections of any of the hypotheses 3 or 4 for calculating the power.
> We now clearly state this in the paper and also clarify that there are different options, and it is up to the user to determine the method of how the power is calculated.

@@

2) There is some inconsistent notation:

a) It seems that n_treat is varyingly defined as the “total number of treatments” (p. 8, l. 29 and p. 9, l. 20), “total number of patients” (p. 8, l. 48 and p. 9, l. 14) and “allowed treatment number” (p. 12, l. 50 and p. 12, l. 51).

> We improved our wording, so that the number of treatment arms cannot longer be confused with the individual number of treatments of each patient.
> Therefore, the total sample size (formerly “total number of treatments”) is denoted as n_total (formerly n_treat).
> Since the patients are different in both stages (i.e. each patient is treated only once) the total number of patients is the same as the total sample size.
> However, we only use the phrase "total sample size" to avoid confusions.

@@

p. 2: For the benefit of readers less familiar with the term, the authors should explain at least briefly what they mean by “expensive black-box” optimisation problems.

> We added a short explanation.

p. 3 ff.: This may be implicitly clear but it should also be said explicitly that in the motivating example and throughout the remainder of the paper a normality assumption is made for all outcome data.

> There is no normality assumption for the individual outcome data but for the test statistics Z_i,j a multivariate normal distribution is assumed which will hold approximately if the sample size is sufficiently large. 
> We clarified that the test statistics are multivariate normal distributed at least approximately at the very end of chapter 3.

@@

p. 11: Whilst I appreciate that the focus is on relative improvements in runtime, the authors should at least briefly describe the hardware setup with which these runtimes were achieved, to provide some context for the absolute runtimes.

> We added the requested information.

p. 12: It is stated that: “Only for the scenarios (effect: paper2, n_treat = 1000) and (effect: sigmoid, n_treat = 1000) Grid Small has a small advantage over BO and Grid.” I think it should be added that Grid Small also has an advantage over BO for (effect: paper, n_treat = 1000) and certainly for (effect: sigmoid, n_treat = 2000).

> Thanks for bringing that up. We revised the list of cases where Grid Small shows a notable advantage.
> An advantage is not notable if the median of both methods lies within the box of the other method.
> Therefore, we do not see a notable advantage of Grid Small for the configuration (effect: paper, n_treat = 1000).

@@

## Reviewer: 2

### Comments to the Author

The authors propose to use Bayesian optimization (BO) to improve the efficiency of the design selection process in clinical trials. A set of parameters to be chose for the design optimization based on the power. The idea seem novel and could be a useful approach. However, there are several major issues and limitations of proposed methods: for example, power is often only part of measure of choosing design, other factors like study duration, and # of pts are also important. Also, the black-box function relay too much of the various parametric assumption of parameters, which need to be carefully decided to be more related to true clinical trial needs. Overall, I have the following comments for the authors.

> We agree that the power is only one measure for choosing a design.
> Therefore, the idea behind our method is to quickly asses which power one can achieve given the other fixed factors such as total sample size (and therefore the number of patients).
> Quickly obtaining the best possible performance for a specific design choice will also help decision makers to decide how to choose the design.
> Furthermore, we added study duration as a possible alternative performance measure. 
> This is now discussed as a limitation in our conclusion section.
> Assuming certain parameters under the null hypothesis is a standard approach.
> However, we see that a more Bayesian approach is certainly interesting, as we already stated in the last paragraph of the discussion.

@@

In section 5.2, the authors keep the total number of pts fixed, which seem not appropriate in practice, as overall power relying several factors, e.g. # of pts, the interim selection rule, stage 1 and stage 2 ratio. Often # of pts is also key goal of optimal design evaluation. I will strongly suggest # of pts should be a parameter in the black-box function, instead of fixing it to a constant value.

> If we included the number of patients in the optimization, the optimizer would propose to choose the maximal number of patients, because this gives the maximal performance.
> This is also indicated in our results, as we tried different values of n_total which equals the number of patients. 
> Here, a higher number always resulted in a higher power.
> As mentioned in the paper, the idea is to run this optimization for different sample sizes (denoted by n_total), and then to decide which sample size is required.
> This brings us to the future goal that we mentioned in the discussion: We can include n_total as another parameter and outcome in a multi-criteria optimization setting.

In page 10 table 4, the number of simulation iteration is only 1000, which does not seem sufficient for complex optimization methods proposed by authors. I would suggest at least 5000 should be needed.

> The simulation iterations are a specific parameter of the black-box function we optimize.
> The higher this value, the more reliable is the estimator of the power of the design.
> Our optimization is run for 100 iterations which is sufficient to converge as Figure 5 shows.
> Therefore, we do not see the need to run a larger number of iterations.
> We added a corresponding statement in section 5.7.

@@

Page 11 table 5, the authors present the average run time in hour, not sure how these information is helpful, as this depend on the what computing systems were used in the simulation. Please either add more information.

> Thank you. We added the hardware setup that was used.

Page 13, BO has issue to find optimal value when one parameters closer to the borders of search space. This is concerning and confirm the limitation of using too many assumptions for optimization, e.g. appropriate parameters range. I am not sure taking log-transform could solve this problem, could authors elaborate more details on this.  Additional simulation is necessary to confirm this.

> As we mentioned in the final discussion section of the paper, we see this as a topic for a future benchmark to further improve the performance of BO.
> The aim of this paper was to benchmark BO in a basic setting without the prior knowledge that small values of **r** are of higher importance.

@@

Given authors used COPD trial as motivating example, I would strongly suggest to add section of case study for the illustration of proposed methods.

> In fact the effect set "paper" is based on a case study where the effect sets are estimated on a real dataset (Friede et al, 2020). 
> In this paper we do not propose a method for analyzing a real dataset, but a method to select the best design given a known theoretical situation.

### 2021-04-12 main.tex
tf notes

BO, often also called Model-based optimization, has been successfully applied to optimize expensive black-boxes (i.e.\ functions that take long to evaluate and are not given in closed-form) in many scenarios.
Popular applications include hyper-parameter tuning of machine learning methods~\citep{snoek_practical_2012}, general algorithm configuration~\citep{hutter_sequential_2011}, and many more.
Various implementations of BO methods are widely used, such as SMAC~\citep{hutter_sequential_2011}, Spearmint~\citep{snoek_practical_2012}, and BoTorch~\citep{balandat_botorch_2020}.
For this work we use the implementation in the R-package \texttt{mlrMBO}~\citep{bischl_mlrmbo_2017}.
This implementation has been successfully used to optimize hyper-parameters of machine learning methods on various tasks~\citep{bischl_mlrmbo_2017, wozniak_candle_2018}, and in the biomedical context it has been applied to optimize model weights~\citep{richter_modelbased_2019,browaeys_nichenet_2020}.
Also, BO is well suited for multi-objective optimization.
Different adaptions of BO exist~\citep{horn_modelbased_2015} that return a set of non-dominated points.
Many of them are implemented in \texttt{mlrMBO} as well.
However, in this work we restrict ourselves to single-objective optimization.

Independently of this work, \citet{wilson_efficient_2021} successfully applied multi-objective BO to minimize the number of participants and the number of clusters in a cluster randomized controlled trial under the restriction that the power is above a given threshold.

@@

One specific trial design that falls into this category are \emph{adaptive seamless designs}~\citep{barnes_integrating_2010} as implemented in the \texttt{asd} package~\citep{parsons_software_2011} and described in Section~\ref{sec:adaptive_seamless_designs}.
%(What is THE asd paper? peter bauer, meinter kieser, 1999 statistics in medicine, % https://doi.org/10.1002/(SICI)1097-0258(19990730)18:14%3C1833::AID-SIM221%3E3.0.CO;2-3, )  bauer_combining_1999
% https://doi.org/10.1002/bimj.200510232 bretz_confirmatory_2006 
% seamless trials (assoziert) 
%  https://doi.org/10.1002/bimj.200510231 schmidli_confirmatory_2006
In~\citet{friede_adaptive_2020} the authors present an application of \emph{adaptive seamless designs} for chronic obstructive pulmonary disease (COPD).
We use this COPD study to apply BO for finding a trial design with maximal statistical power, given the parameters of the respective design options.
%within the given scenarios restrictions.

@@

The evaluation of some performance measures requires Monte Carlo simulations, since no closed form solutions are available. 
The R-package \texttt{asd} by~\citet{parsons_software_2011} can be used for such simulations.
As it generates test statistics, following multivariate normal distributions, it is applicable to a wide range of outcomes, at least approximately, and generally more efficient in comparison to simulating individual participant data.
The simulation model was described in more detail by \citet{friede_adaptive_2020}. 
It features selection at interim based on early as well as final outcomes.

@@

The expected improvement, as well as other acquisition functions, are derived under the assumption that at each point $\theta \in \Theta$ the posterior of the function outcome follows a normal distribution.
The standard Gaussian process regression, that is often used as surrogate, models the posterior as a Gaussian with the parameters $\hat{\mu}$ and $s^2(\theta)$ under the assumption that $f$ is a realization of a Gaussian process.
In practice, this assumption often cannot be verified.
However, benchmarks show that the approach nevertheless works reliable for many practical problems~\citep{bischl_mlrmbo_2017,snoek_practical_2012}.

@@

The goal of this simulation study is to investigate the potential benefit of BO for efficiently finding a clinical trial design that maximizes a given metric.
The analysis is motivated by the clinical example with COPD patients introduced in Section~\ref{sec:motivating_example}, with treatment selection in the first of two stages.
In the first stage five arms (defined by the control and four different administered doses) are considered, and in the second stage only a selected subset of arms is kept for further evaluation. 
The metric to be optimized is the statistical power, where a positive outcome means a trial arm with true effect is detected as significant effect.
Depending on the application the power can be calculated as the rate of rightfully rejecting any, all, or a subset of the elementary hypotheses~\citet{senn_power_2007}.

@@

\paragraph{Grid and Grid small}
We conduct an \emph{exhaustive grid search} with a resolution of $l$ points per real-valued dimension and one point for each categorical dimension.
Referring to our application, for each selection strategy we evaluate a numerical grid with a resolution of $l$.
All 6 selection strategies have the numeric parameter $r$.
The strategies \emph{epsilon} and \emph{thresh} both have an additional second parameter, $\epsilon$ or $\tau$, respectively.
For $l=25$, we call this approach \textbf{Grid},
with a total of $4 \cdot 25 + 2 \cdot 25 \cdot 25 = 1350$ design configurations that are evaluated for one scenario.
We chose $l=25$ generously to obtain an estimate of the optimum that should be close to the true optimum.
Previous studies used half the resolution to assess the behavior of similar problems~\citep{friede_adaptive_2020}.
We define \textbf{Grid Small} to be a subset of this grid with $l=7$ points per dimension, resulting in $4 \cdot 7 + 2 \cdot 7 \cdot 7 = 126$.
This gives a number of evaluations that can be easily carried out in practice.

@@

As an acquisition function we choose the $\operatorname{AEI}$ as explained in Section~\ref{sec:bayesian_optimization}.
We start BO with an initial design of $16$ randomly sampled points, following an established rule-of-thumb of $4$ points per dimension, which is also well in line with the recommendation of using around $10\%$ of the budget for the initial design~\cite{bossek_initial_2020}.
We allow $100$ further iterations, summing up to a total budget of $116$ evaluated configurations of the simulator $\tilde{f}$.

@@

The algorithms used in this study are implemented in R.
For the Bayesian optimization, the R-package \texttt{mlrMBO}~\citep{bischl_mlrmbo_2017} is used.
For the simulation of clinical trials, the function \texttt{treatsel.sim} of the R-package \texttt{asd}~\citep{parsons_software_2011} is used, with parameter values given in Table~\ref{tab:par_implement}.
%The vector of treatment numbers for determining power \texttt{ptest} counts rejections of treat3 or 4 hypotheses. %???
The vector of treatment numbers \texttt{ptest} determines the hypotheses for which rejections are counted towards the power.
Here, the rejection of one or both of the hypotheses that treatment 3 or 4 has no significant effect against the control determines the power, as in~\citet{friede_adaptive_2020}.

@@

Figure~\ref{fig:plot_opt_path_5000} shows that indeed the lower noise level leads to more stable simulation runs, meaning that single optimization runs lead to results closer to the mean result.
Both curves indicate, that more iterations of \emph{BO} would not yield substantially better outcomes as the progress is stagnating after 80 iterations.
It has to be noted that this plot reports the best observed outcome so far and not the one that is chosen as final result.
%So on the left side of Figure~\ref{fig:plot_opt_path_5000}, we see more curves that observed overoptimistic outcomes in comparison the the right side.
%Ideally, the choice of the final point should not be influenced by these overoptimistic observations for \emph{BO} as explained in Section~\ref{ssec:best_point}, as the final point is determined by the surrogates mean prediction.
Despite the differences of the optimization paths, the median validation error reported in Figure~\ref{fig:plot_boxplot_valid_y_5000} is nearly identical for \emph{BO} for the low and the high value of $n_\text{sim}$.
Therefore, we can assume that "bias correction" of \emph{BO} as explained in Section~\ref{ssec:best_point} seems to be effective.
\begin{figure}[htb]
\centering
  \begin{minipage}{0.62\textwidth}
    \centering
    \includegraphics[width=\linewidth]{generated/figures/plot_opt_path_5000.pdf}%
    \captionof{figure}{Optimization progress for \emph{BO}: Best outcome found so far for each iteration, in comparison to the best outcome found in the grid search, for effect set \emph{paper} and $n_{\text{total}} = 2000$}
    \label{fig:plot_opt_path_5000}
  \end{minipage}\hspace{0.02\textwidth}%
  \begin{minipage}{0.35\textwidth}
    \centering
    \includegraphics[width=\linewidth]{generated/figures/plot_boxplot_valid_y_5000.pdf}%
    \captionof{figure}{Validated performance of all optimization strategies on effect set \emph{paper} and $n_{\text{total}} = 2000$ with 100 stochastic repetitions and different values for \texttt{nsim}}
    \label{fig:plot_boxplot_valid_y_5000}
  \end{minipage}
\end{figure}
For all methods the IQR displayed in the Box plots is much smaller for $n_\text{sim} = 5000$, indicating that the solutions that are found vary less between the 100 stochastic replications.
Also, the absolute differences between the methods are even smaller now.
This demonstrates that for the selected scenario all methods can find the optimal configuration if the results are not noisy.
However, an increased value of $n_\text{sim}$ also increases the runtime linearly, which conflicts with the goal to find the optimal configuration quickly.
In summary, \emph{BO} can find configurations with a small number of required simulations (higher noise) that are nearly as good as the optimal solution which is identified with a large number of simulations.

@@

% subsection Summary
Bayesian optimization (BO) is a state-of-the art optimizer that can be used for parameter optimization in a wide range of applications.
Here, we propose, for the first time, to optimize the power of adaptive seamless designs using BO.
In clinical trials, determining the test methods and sample sizes is a key step in the planning process.
One main goal is to maximize the test power, given a set of treatments with corresponding effect sizes and a fixed number of samples.
If determining the test power involves intensive simulations, trying out different treatment design configurations can become a laborious task.
Either large computational resources are required for an exhaustive investigation of all the different configurations, or a tedious manual trial-and-error process is required.

@@

% subsection Outlook
Only in cases in which the optimal values of one parameter are close to the lower border, we sometimes observed poor performance of BO.
Here, a possible improvement can be to apply a log transformation to such parameters. % in the future.
Further, instead of keeping the total sample size constant, we could apply multi-objective optimization and include this number as an additional outcome.
This multi-objective approach would result in a set of points (a Pareto front) that minimizes the total sample size and at the same time maximize the statistical power of the trial design.
The user could then see for which number of trials the power is above a certain threshold (e.g. $95\%$).
Another way to speed up the optimization would be to iteratively increase the number of simulation rounds of the trial design evaluation. 
If few simulations already indicate a bad outcome for a configuration then additional simulation rounds are not necessary.
Such a procedure can be guided by statistical testing and is already common in machine learning under the terms \emph{racing} or \emph{multi-fidelity optimization}.

>>>>

BO, often also called Model-based optimization, has been successfully applied to optimize expensive black-boxes (i.e.\ functions that take long to evaluate and are not given in closed-form) in many scenarios.
Popular applications include hyper-parameter tuning of machine learning methods~\citep{snoek_practical_2012}, general algorithm configuration~\citep{hutter_sequential_2011}, and many more.
Various implementations of BO methods are widely used, such as SMAC~\citep{hutter_sequential_2011}, Spearmint~\citep{snoek_practical_2012}, and BoTorch~\citep{balandat_botorch_2020}.
For this work we use the implementation in the R-package \texttt{mlrMBO}~\citep{bischl_mlrmbo_2017}.
This implementation has been successfully used to optimize hyper-parameters of machine learning methods on various tasks~\citep{bischl_mlrmbo_2017, wozniak_candle_2018}, and in the biomedical context it has been applied to optimize model weights~\citep{richter_modelbased_2019,browaeys_nichenet_2020}.
Also, BO is well suited for multi-objective optimization.
Different adaptions of BO exist~\citep{horn_modelbased_2015} that return a set of non-dominated points.
Many of them are implemented in \texttt{mlrMBO} as well.
However, in this work we restrict ourselves to single-objective optimization. 
Independently of this work, for instance, \citet{wilson_efficient_2021} successfully applied multi-objective BO to minimize the number of participants and the number of clusters in a cluster randomized controlled trial under the restriction that the power is above a given threshold.

@@

One specific trial design that falls into this category are \emph{adaptive seamless designs}~\citep{barnes_integrating_2010} as implemented in the \texttt{asd} package~\citep{parsons_package_2012} and described in Section~\ref{sec:adaptive_seamless_designs}.
%(What is THE asd paper? peter bauer, meinter kieser, 1999 statistics in medicine, % https://doi.org/10.1002/(SICI)1097-0258(19990730)18:14%3C1833::AID-SIM221%3E3.0.CO;2-3, )  bauer_combining_1999
% https://doi.org/10.1002/bimj.200510232 bretz_confirmatory_2006 
% seamless trials (assoziert) 
%  https://doi.org/10.1002/bimj.200510231 schmidli_confirmatory_2006
In~\citet{friede_adaptive_2020} the authors present an application of \emph{adaptive seamless designs} for chronic obstructive pulmonary disease (COPD).
We use this COPD study to apply BO for finding a trial design with maximal statistical power, given the parameters of the respective design options.
%within the given scenarios restrictions.

@@

The evaluation of some performance measures requires Monte Carlo simulations, since no closed form solutions are available. 
The R-package \texttt{asd} by~\citet{parsons_package_2012} can be used for such simulations.
As it generates test statistics, following multivariate normal distributions, it is applicable to a wide range of outcomes, at least approximately, and generally more efficient in comparison to simulating individual participant data.
The simulation model was described in more detail by \citet{friede_adaptive_2020}. 
It features selection at interim based on early as well as final outcomes.

@@

The expected improvement as well as other acquisition functions are derived under the assumption that at each point $\theta \in \Theta$ the posterior of the function outcome follows a normal distribution.
The standard Gaussian process regression, that is often used as surrogate, models the posterior as a Gaussian with the parameters $\hat{\mu}$ and $s^2(\theta)$ under the assumption that $f$ is a realization of a Gaussian process.
In practice, this assumption often cannot be verified.
However, benchmarks show that the approach nevertheless works reliable for many practical problems~\citep{bischl_mlrmbo_2017,snoek_practical_2012}.

@@

The goal of this simulation study is to investigate the potential benefit of BO for efficiently finding a clinical trial design that maximizes a given metric.
The analysis is motivated by the clinical example with COPD patients introduced in Section~\ref{sec:motivating_example}, with treatment selection in the first of two stages.
In the first stage five arms (defined by the control and four different administered doses) are considered, and in the second stage only a selected subset of arms is kept for further evaluation. 
The metric to be optimized is the statistical power, where a positive outcome means a trial arm with true effect is detected as significant effect.
Depending on the application the power can be calculated as the proportion of correctly rejecting any, all, or a subset of the elementary hypotheses~\citep{senn_power_2007}.

@@

\paragraph{Grid and Grid small}
We conduct an \emph{exhaustive grid search} with a resolution of $l$ points per real-valued dimension and one point for each categorical dimension.
Referring to our application, for each selection strategy we evaluate a numerical grid with a resolution of $l$.
All 6 selection strategies have the numeric parameter $r$.
The strategies \emph{epsilon} and \emph{thresh} both have an additional second parameter, $\epsilon$ or $\tau$, respectively.
For $l=25$, we call this approach \textbf{Grid},
with a total of $4 \cdot 25 + 2 \cdot 25 \cdot 25 = 1350$ design configurations that are evaluated for one scenario.
We chose $l=25$ generously to obtain an estimate of the optimum that should be close to the true optimum.
Previous studies used half the resolution to assess the behavior of similar problems~\citep{friede_adaptive_2020}.
We define \textbf{Grid Small} to be a subset of this grid with $l=7$ points per dimension, resulting in $4 \cdot 7 + 2 \cdot 7 \cdot 7 = 126$.
This results in a number of evaluations that can be easily carried out in practice.

@@

As an acquisition function we choose the $\operatorname{AEI}$ as explained in Section~\ref{sec:bayesian_optimization}.
We start BO with an initial design of $16$ randomly sampled points, following an established rule-of-thumb of $4$ points per dimension, which is also well in line with the recommendation of using around $10\%$ of the budget for the initial design~\citep{bossek_initial_2020}.
We allow $100$ further iterations, summing up to a total budget of $116$ evaluated configurations of the simulator $\tilde{f}$.

@@

The algorithms used in this study are implemented in R.
For the Bayesian optimization, the R-package \texttt{mlrMBO}~\citep{bischl_mlrmbo_2017} is used.
For the simulation of clinical trials, the function \texttt{treatsel.sim} of the R-package \texttt{asd}~\citep{parsons_package_2012} is used, with parameter values given in Table~\ref{tab:par_implement}.
%The vector of treatment numbers for determining power \texttt{ptest} counts rejections of treat3 or 4 hypotheses. %???
The vector of treatment numbers \texttt{ptest} determines the hypotheses for which rejections are counted towards the power.
Here, the rejection of one or both of the hypotheses that treatment 3 or 4 has no effect against the control determines the power, as in~\citet{friede_adaptive_2020}.

@@

Figure~\ref{fig:plot_opt_path_5000} shows that indeed the lower noise level leads to more stable simulation runs, meaning that single optimization runs lead to results closer to the mean result.
Both curves indicate, that more iterations of \emph{BO} would not yield substantially better outcomes as no substantial improvements are observed after 80 iterations
It has to be noted that this plot reports the best observed outcome so far and not the one that is chosen as final result.
%So on the left side of Figure~\ref{fig:plot_opt_path_5000}, we see more curves that observed overoptimistic outcomes in comparison the the right side.
%Ideally, the choice of the final point should not be influenced by these overoptimistic observations for \emph{BO} as explained in Section~\ref{ssec:best_point}, as the final point is determined by the surrogates mean prediction.
Despite the differences of the optimization paths, the median validation error reported in Figure~\ref{fig:plot_boxplot_valid_y_5000} is nearly identical for \emph{BO} for the low and the high value of $n_\text{sim}$.
Therefore, we can assume that "bias correction" of \emph{BO} as explained in Section~\ref{ssec:best_point} seems to be effective.
\begin{figure}[htb]
\centering
  \begin{minipage}{0.62\textwidth}
    \centering
    \includegraphics[width=\linewidth]{generated/figures/plot_opt_path_5000.pdf}%
    \captionof{figure}{Optimization progress for \emph{BO}: Best outcome found so far for each iteration, in comparison to the best outcome found in the grid search, for effect set \emph{paper} and $n_{\text{total}} = 2000$}
    \label{fig:plot_opt_path_5000}
  \end{minipage}\hspace{0.02\textwidth}%
  \begin{minipage}{0.35\textwidth}
    \centering
    \includegraphics[width=\linewidth]{generated/figures/plot_boxplot_valid_y_5000.pdf}%
    \captionof{figure}{Validated performance of all optimization strategies on effect set \emph{paper} and $n_{\text{total}} = 2000$ with 100 stochastic repetitions and different values for \texttt{nsim}}
    \label{fig:plot_boxplot_valid_y_5000}
  \end{minipage}
\end{figure}
For all methods the IQR displayed in the Box plots is much smaller for $n_\text{sim} = 5000$, indicating that the solutions that are found vary less between the 100 stochastic replications.
Also, the absolute differences between the methods are even smaller now.
This demonstrates that for the selected scenario all methods can find the optimal configuration if the results are not noisy.
However, an increased value of $n_\text{sim}$ also increases the runtime linearly, which conflicts with the goal to find the optimal configuration quickly.
In summary, \emph{BO} can find configurations with a small number of required simulations (higher noise) that are nearly as good as the optimal solution which is identified with a large number of simulations.

@@

% subsection Summary
Bayesian optimization (BO) is a state-of-the art optimizer that can be used for parameter optimization in a wide range of applications.
Here, we propose, for the first time, to optimize the power of adaptive seamless designs using BO.
In clinical trials, determining the test methods and sample sizes is a key step in the planning process.
For our optimization we chose to maximize the test power, given a set of treatments with corresponding effect sizes and a fixed number of samples.
%FIXME: TF: Sollten wir hier nochmal explizit auf alternative objectives wie study duration eingehen?
If determining the test power involves intensive simulations, trying out different treatment design configurations can become a laborious task.
Either large computational resources are required for an exhaustive investigation of all the different configurations, or a tedious manual trial-and-error process is required.

@@

% subsection Outlook
Only in cases in which the optimal values of one parameter are close to the lower border, we sometimes observed poor performance of BO.
Here, a possible improvement can be to apply a log transformation to such parameters. % in the future.
We chose to maximize the power for a fixed number of samples.
Other aspects like study duration or the cost of a trial in general could be of interest either as an additional constraint or an additional outcome.
Also the number of samples can be included as an additional outcome instead of being fixed.
Multiple outcomes that are subject to optimization give a (possibly restricted) multi-objective optimization problem.
The result of which is a set of points (a Pareto front) that maximizes the statistical power and e.g. minimizes the total sample size at the same time.
The user could then see for which sample size the power is above a certain threshold (e.g. $95\%$).
Another way to speed up the optimization would be to iteratively increase the number of simulation rounds of the trial design evaluation. 
If few simulations already indicate a bad outcome for a configuration then additional simulation rounds are not necessary.
Such a procedure can be guided by statistical testing and is already common in machine learning under the terms \emph{racing} or \emph{multi-fidelity optimization}.
